Predicting Spotify Audio Features from Last.fm Tags

### 2023-06-03 paper/main.tex
draft experiments results

The preceding models require specific input formats.
Therefore, we tested different input formats.

Each individual sample in the dataset corresponds to a unique track,
and contains the list of Last.fm tag-count tuples (e.g. \verb|[(electronic, 100), (dance, 45), ...]|)
and the values of Spotify audio features.
Before experimenting with machine learning models, we prepared the data in a number of different formats,
each one suitable for specific models.

\subsubsection{Tabular}
With this format, the Last.fm tags are represented as a table.
Each tag is defined by a column and each cell contains the count value of a tag for a track.
A cell is 0 if a tag is missing for a track.

Counting the total amount of Last.fm tags in the user collection resulted, initially, in more than five million tags.
We quickly confirmed that building a tabular data set, in which every row contains millions of columns (Last.fm tags)
was doable, but presented scalability problems.
In addition to scalability limitations, classic machine learning models might not take advantage of using the full dataset.
These models might even perform poorly if too many input features are provided.
The reason for this is spurious relations or redundancy  between input features.
Models might find relations that are not real, and latent, redundant variables might be accountable multiple times, which creates a bias in the output.
Feature subject selection aims at solving these problems.

% For example, a Bayesian model might provide low-quality results of provided with variables that are not conditionally independent.

Therefore, we reduced the number of tags by picking a subset of the most relevant tags.

The reduction algorithm uses a basic data aggregation algorightm: group the data by tag, aggregate by summing the \emph{count} values for each tag,
order by the aggregated count, and finally select the top-$K$ items of the ranking.

We generated three versions of the dataset, with different values of $K$: 100, 1,000 and 10,000.

We consider this reduction approach an initial approach.
For this particular aspect, dimensionality reduction algorithms, such as PCA, are good candidates for future work.

After selecting the top-$K$ tags, the input data passed was formatted in tabular format, as follows:

\begin{itemize}
      \item Given that $Tags_{K}$ is the set of most $K$ frequent Last.fm tags in the user listening history
            and, where each $tag \in Tags_{K}$.
      \item Given that $Audio$ is the set of Spotify audio features, where each $feat \in Audio$.
      \item For each $track$:
      \begin{itemize}
            \item $X_{track},_{tag}$ is the strengh of $tag$ for $track$. This value is in the $0-100$ range.
            \item $y_{track},_{feature}$ is the value of the audio feature $y$ for $track$.
      \end{itemize}
\end{itemize}

@@

When generating training data by track, the tabular formats present sparsity problems.

For tabular representations, we need to defined a fixed set of columns as tags.
For most of tracks, most columns are 0.

The sparsity of a matrix is the number of zero-valued elements divided by the total number of elements
(e.g., m * n for an m * n matrix) is called the sparsity of the matrix.

\subsubsection{Tabular Tokens}

Tags are converted to text tokens. Columns represent token positions, and cells contain the token at a particular position, for a track.
To tokenize tags, we have used the GTP-2 tokenizer.

Tokenizers are crucial elements in the preprocessing of text data.
A tokenizer dissects a piece of text into smaller units, called tokens.
These tokens can be words, subwords, or even syllables or characters.

When breaking down the text into tokens, the tokenizer assigns a unique numerical identifier to each token.
This IDs are based on the vocabulary that the tokenizer has been trained on.
For example, when the tokenizer processes the \verb|"Hello world"| string, the Hello token might be assigned to ID \verb|123|
and the world token might be assigned to ID \verb|34534|.
The result of tokenizing \verb|Hello world| would be \verb|[123, 34534]|.

Note that, although tokenizers are most commonly used in combination with transformer models, in this paper we test the possiblity of using theorem
to preproccess data passed to Bayesian an Tree models.

Because the tokenizer requires a string as input, we have converted the set of tags for each track into a string.
To \emph{stringify} the tags, we have concatenated tags with multiple strategies:

\begin{itemize}
      \item Ordering by count: \verb|"rock, pop"|.
      \item Including tag count: \verb|"rock 2, pop 1"|.
      \item Duplicating each tag $count$ times: \verb|"rock rock, pop"|.
\end{itemize}

In this particular case, the $X$ values of the tabular input data are tokens.
These tokens are obtained from passing the string of concatenated Last.fm tags through the GPT-2 tokenizer,
as the following procedure explains:

\begin{itemize}
      \item Given that $X_L$ is the token vocabulary, where $L$ is the maximum vocabulary length.
      \item Given that $Audio$ is the set of Spotify audio features, where each $feat \in Audio$.
      \item For each $track$:
      \begin{itemize}
            \item $X_{track},_{n}$ is token found at position $n$, after tokenizing the tags string.
            \item $y_{track},_{feature}$ is the value of the audio feature $y$ for $track$.
      \end{itemize}
\end{itemize}

@@

Similarly to the tags tabular format, we also defined fixed values for the number of columns: 10, 1,000, and 10,000.

The two tabular formats, with tags and tokens, were tested on the Naive Bayes and the Boosted tree regressors.

\subsection{Text Strings}

When using transformer models, the input data is a string.
We must represent the Last.fm tags, which are defined as $(tag, count)$ tuples, as strings.
To this end, we applied the same three transformations used in the tabular tokens formats,
concatenate the tags by ordering by count, by including the tag count, and by duplicating the tags.

@@

@@

      Tree      & 10000 tags                          &\textbf{0.148}    & 0.250           &\textbf{0.181}  &\textbf{0.235}   &\textbf{0.290}  \\

@@

      Tree      & 1000 tokens D                       & 0.173            & 0.278           & 0.215          & 0.220           & 0.339          \\

@@

      Tree      & 1000 tokens TC                      & 0.174            & 0.282           & 0.215          & 0.220           & 0.344          \\

@@

@@

>>>>

The preceding models require specific input formats.
In particular, the boosted tree and bayesian ridge regressors require structured data (e.g a set of predictor features and a set of target variables), whereas GTP-2 expects text strings as input.

Each individual sample in the Last.fm singe-user dataset corresponds to a unique track,
and contains the list of Last.fm tag-count tuples (e.g. \verb|[(electronic, 100), (dance, 45), ...]|)
and the values of Spotify audio features.
The Last.fm single-user dataset was converted to the formats described in the following sections, and then fed into the corresponding models.

\subsubsection{Last.fm Tags as Table Columns}
The tabular format represents the Last.fm tags as columns in a table.
Each tag is defined by a column and each cell contains the count value of a tag for a track.
If a $tag$ is not present for a particular $track$, then cell $c_{track,tag}$ is \num{0}.

Counting the total amount of Last.fm tags in the user collection resulted, initially, in more than five million tags.
Under this high-dimensionality scenario, building a tabular data set, in which every row contains millions of columns (i.e. Last.fm tags) was theoretically possible, but presented scalability problems.
In addition to scalability limitations, classic machine learning models might not take advantage of using the full collection of tags present in the user history.
These models might even perform poorly if too many input features are provided.
The reason for this is spurious relations or redundancy  between input features.
Models might find relations that are not real, and latent, redundant variables might be accountable multiple times, which can lead to a biased outputs.
Feature subject selection aims at solving these problems.

Therefore, we reduced the number of tags by picking a subset of the most relevant tags.
We used a the feature selection algorithm by using a basic data aggregation algorithm: grouping the data by tag, aggregating by summing the \emph{count} values for each tag,
ordering by the aggregated count, and finally selecting the top-$K$ tags of the ranking.
Three values of $K$, were used, thus generating three subsets: \num{100}, \num{1000}, and \num{10000}.
We consider this reduction approach an initial approach approach in our experiments.
For this particular aspect, dimensionality reduction algorithms, such as PCA, are good candidates for future work.

After selecting the top-$K$ tags, the Last.fm single-user dataset was formatted as follows:

\begin{itemize}
      \item Given that $Tags_{K}$ is the set of most $K$ frequent Last.fm tags in the user listening history
            , where each $tag \in Tags_{K}$.
      \item Given that $Audio$ is the set of Spotify audio features, where each $feat \in Audio$.
      \item For each $track$:
      \begin{itemize}
            \item $X_{tag,track}$ is the $count$ value of $tag$ for $track$. This value is in the $0-100$ range.
            \item $y_{track},_{feature}$ is the value of the audio feature $y$ for $track$.
      \end{itemize}
\end{itemize}

@@

The main drawback of this representation is data sparsity.
For most of the tracks, many columns are \num{0}.

This format was tested with the Bayesian Ridge and the Boosted Tree regressors.

% The sparsity of a matrix is the number of zero-valued elements divided by the total number of elements
% (e.g., m * n for an m * n matrix) is called the sparsity of the matrix.

\subsubsection{Tokens as Table Columns}

This format is also structured, but the input data is a set of tokens instead of tag count values.
These \emph{tokens} are the result of passing the tags, concatenated as a string, through a tokenizer.
Tokenizers are crucial elements in the preprocessing of text data.
A tokenizer dissects a piece of text into smaller units, called tokens.
These tokens can be words, subwords, or even characters.

% Tags are converted to text tokens. Columns represent token positions, and cells contain the token at a particular position, for a track.
% To tokenize tags, we have used the GTP-2 tokenizer.

When breaking down the text into tokens, the tokenizer assigns a unique numerical identifier to each token.
These identifiers are based on the vocabulary that the tokenizer has been trained on.
For example, when the tokenizer processes the \verb|"pop rock"| string, the \verb|pop| token might be assigned to ID \verb|123|
and the \verb|rock| token might be assigned to ID \verb|34534|.
Consequently, the result of tokenizing \verb|pop rock| would be \verb|[123, 34534]|.

Note that, although tokenizers are most commonly used in combination with transformer models, in this paper we test the possibility of using a tokenizer to preproccess the data passed to the models that require structured data.

Because the tokenizer requires a string as the input, we converted the set of tags for each track into a string, in a process that we called \emph{stringification}.
To \emph{stringify} the tags, we concatenated Last.fm tags by following these three strategies:

\begin{itemize}
      \item Order by count: \verb|"rock, pop"|.
      \item Include tag count: \verb|"rock 2, pop 1"|.
      \item Duplicate tags $count$ times: \verb|"rock rock, pop"|.
\end{itemize}

In this particular case, the $X$ values of the tabular input data are the token IDs.
These tokens are obtained from passing the string of concatenated Last.fm tags through the GPT-2 tokenizer,
as the following procedure explains:

\begin{itemize}
      \item Given that $S$ is the stringification strategy.
      \item Given that $X_L$ is the token vocabulary, where $L$ is the maximum vocabulary length.
      \item Given that $Audio$ is the set of Spotify audio features, where each $feat \in Audio$.
      \item For each $track$ and $S$:
      \begin{itemize}
            \item $tags_{track,str}$ is the string of concatenated tags produced by strategy $S$.
            \item $tokens_{track}$ is the list of token IDs produced after tokenizing $tags_{str}$.
            \item $X_{n,track}$ is token ID found at position $n$ of $tokens_{track}$.
            \item $y_{feature, track}$ is the value of the audio feature $y$ for $track$.
      \end{itemize}
\end{itemize}

@@

Similarly to the tags tabular format, we also defined fixed values for the number of columns: 10, 1,000, and 10,000.
\textcolor{red}{Decir cual es el nombre de este limite (maxsequence length?)}

Similar to the previous structured format, this format was also used in the Bayesian and Boosted Tree regressors.
In contrast to the previous format, this format does not present sparsity.
It incorporates however, other problems, such as the lack of normalization of the values, which do not represent scalars, but IDs.
It is also harder to inspect, because the meaning of each token is unknown.

\subsubsection{Text Strings}

When using transformer models, the input data is a string.
To use GPT-2, we had to represent the Last.fm tags, which are defined as $(tag, count)$ tuples, as strings.
To this end, we applied the same three transformations used in the tabular tokens formats,
concatenate the tags by ordering by count, by including the tag count, and by duplicating the tags.

@@

The experiments tested the three models, and the three values of $K$: \num{100}, \num{1000}, \num{10000}.
In the Bayesian Ridge and Boosted tree regressors, we tested the two structured data formats: tags and tokens as columns.
In the GTP-2 model, we tested the text data format.

For the formats that required the concatenation of tags into text strings, we tested the three different stringification strategies.

The data was split into a training set (\num{8654} samples), a validation set(\num{2164} samples), and a test set (\num{2705} samples).
The quality of the model was evaluated by using the root mean squared error (RMSE).
This metric was computed on the test set.

@@

      Tree      & 10000 tags                          &\textbf{0.148}    & 0.250           &\textbf{0.181}  & 0.235           &\textbf{0.290}  \\

@@

      Tree      & 1000 tokens D                       & 0.173            & 0.278           & 0.215          & \textbf{0.220}  & 0.339          \\

@@

      Tree      & 1000 tokens TC                      & 0.174            & 0.282           & 0.215          & \textbf{0.220}  & 0.344          \\

@@

The Spotify feature that the models had less trouble estimating was the danceability.
The instrumentalness, however, was the feature that presented the highest deviations.
The GTP-2 model was the best model to predict the danceability, with a \num{0.145} RMSE.
The Boosted tree was the best model to predict the instrumentalness, with a RMSE value of \num{0.29}.

In general, the three models performed similarly, with GPT-2 achieving the slightly better results in danceability and acousticness. The Boosted Tree regressor was the best model to predict energy, valence, and instrumentalness.

The two regressors achieved the best results when using the \emph{Tags as columns} format, mostly with \num{10000} tags.
Interestingly, the Boosted Tree regressor achieved better results with tokens when estimating valence.
Even better results than GPT-2.
In the rest of the features, the regressors worked better with tags as columns.

Using a higher number of predictor variables was not a synonym of lower error values when using tokens a columns.

The Bayesian regressor was a bit more sensitive to changes in tag formating.
This model tended to work better with data formatted by using the Tag Count stringification method.
The Boosted Tree regressor presented less sensitivity to format changes, although the model generally performed better with the Tags as columns format too.

The Duplicated stringification method consistently performed worse with the GTP-2 transformer model.

@@

We consider this reduction approach an initial approach approach in our experiments.
For this particular aspect, dimensionality reduction algorithms, such as PCA, are good candidates for future work.
